---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(Epi)
library(kableExtra)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r, collapse=T}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Write functions that describes the data

```{r}

## Identify the cohort that entered the trial

trial_cohort <- function(dataset){
  
  trial <- dataset %>% 
    subset(trial_participant == TRUE) %>% 
    # Identify the observed outcomes
    mutate(preg_outcome_final = ifelse(trt == 1,
                                       preg_outcome_final_trt,
                                       preg_outcome_final_untrt),
           preg_outcome_final_gw = ifelse(trt == 1,
                                          preg_outcome_final_gw_trt,
                                          preg_outcome_final_gw_untrt),
           preeclampsia = ifelse(trt == 1,
                                 preeclampsia_trt,
                                 preeclampsia_untrt),
           sga = ifelse(trt == 1,
                        sga_trt,
                        sga_untrt)) %>% 
    rowwise() %>% 
    mutate(censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                      1,
                      0),
           time = ifelse(censor == 1,
                         first_censoring_gw,
                         preg_outcome_final_gw)
           ) %>% 
    ungroup() %>% 
    mutate(
      composite_km = case_when(censor == 1 ~ 0,
                          preg_outcome_final == "fetaldeath" & censor == 0 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 1 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 0 ~ 0),
      composite_multi = case_when(censor == 1 ~ 0,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 1,
                                censor == 0 & preeclampsia == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 2
                                ),
      # Make a gestational week variable for the composite_multi variable
      composite_multi_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # Make composite variable that ignores censoring
      composite_no_censor = ifelse(preg_outcome_final == "fetaldeath" | 
                                     (preg_outcome_final == "livebirth" & preeclampsia == 1),
                                   1,
                                   0),
      # Make a SGA variable for the KM estimator
      sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0),
      # Make SGA variables with fetal death only as a competing event
      sga_multi_do = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 0), 
      # Make a gestational week variable for the sga_multi variable
      sga_multi_do_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # Make SGA variables with all competing events
      sga_multi_all = case_when(censor == 1 ~ 0,
                                censor == 0 & sga == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                                censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 2), 
      # COULD BE 3 - COME BACK
      # Make a gestational week variable for the sga_multi variable
      sga_multi_all_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # Make a SGA variable that ignores censoring
      sga_no_censor = ifelse(sga == 1, 1, 0),
      
      # PReterm KM variables
      preterm_km = ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
                          preg_outcome_final == "livebirth", # Previously 37
                          1,
                          0),
      preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
                         ifelse(preterm_km == 0 & time >= 35, 35,
                                time)),
      # Preterm AJ variables - NEED TO CHECK
      preterm_multi = case_when(time >= 35 ~ 0,
                                censor == 1 & time < 35 ~ 0,
                                censor == 0 & preg_outcome_final == "livebirth" & time < 35 ~ 1,
                                censor == 0 & preg_outcome_final == "fetaldeath" & time < 35 ~ 2),
      preterm_multi_gw = ifelse(time >= 35, 35, time),
      # Make a preterm variable that ignores censoring
      preterm_no_censor = ifelse(preg_outcome_final == "livebirth" & preg_outcome_final_gw < 35,
                                 1,
                                 0)

    )
  
  return(trial)
  
}

    


### Calculate the number of trial participants

total_trial <- function(dataset){
  
  tib <- dataset %>% 
    summarize(count = n()) 

  return(as.double(tib[1,1]))
  
}

## Count the number of patients across treatment groups
## -- Will save this as a tibble

total_trial_trt <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(count = n(),
              percent = round(count / nrow(dataset) * 100),2)
  
  return(tib)
  
}

## Describe indexing into the cohort

trial_index <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(
              min = min(first_index),
              p25 = quantile(first_index, 0.25),
              median = median(first_index),
              p75 = quantile (first_index, 0.75),
              max = max(first_index),
              nmiss = sum(is.na(first_index))
              )
  
  return(tib)
  
}

## Number of live births and fetal deaths across treatment groups
## -- Currently, this does not consider censoring.
count_outcomes <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
              fd_percent = round(fetal_death / n() * 100,1),
              livebirth = sum(preg_outcome_final == 'livebirth'),
              lb_percent = round(livebirth / n() * 100,1))
  
  return(tib)
  
}

## Distribution of pregnancy outcomes
## -- Currently, this does not consider censoring.
outcomes_dist <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt, preg_outcome_final) %>% 
    summarize(
              min = min(preg_outcome_final_gw),
              p25 = quantile(preg_outcome_final_gw, 0.25),
              median = median(preg_outcome_final_gw),
              p75 = quantile (preg_outcome_final_gw, 0.75),
              max = max(preg_outcome_final_gw),
              .groups = 'keep'
              )
  
  return(tib)
  
}

## Number with preeclampsia and SGA
count_p_s <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(preeclampsia = sum(preeclampsia == 1),
              preec_perc = round(preeclampsia / n() * 100, 1),
              sga = sum(sga == 1),
              sga_perc = round(sga / n() * 100, 1))
  
  return(tib)
  
}

## KM Estimator
## -- Code for function that uses the Kaplan-Meier estimator to estimate risk.

km_estimator <- function(dataset, outcome_var, outcome_var_t, t_val){
  
  ## Run the KM analysis only considering the composite outcome.
  km <- survfit(Surv(get(outcome_var_t), get(outcome_var)) ~ trt, data = dataset)
  
  wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, #se = km$std.err,
                          trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))
  
  estimate <- wrisk %>% 
    filter(t < t_val+0.5) %>% # Deal with no additional risk scenarios
    group_by(trt) %>% 
    summarize(r = last(r),
              #se = last(se),
              .groups = 'drop') %>% 
    pivot_wider(names_from = trt,
                values_from = c(r), #, se
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           Estimator = "Kaplan-Meier") %>%
    select(r_0, r_1, rr, #RR_lower, RR_upper, 
           rd #, 
           #RD_lower, RD_upper
           )

  return(estimate)
  
}


## AJ Estimator

aj_estimator <- function(dataset, outcome_var, outcome_var_t, t_val){
  
  #jitter ties
  #set.seed(1234)
  # Not sure if should set seed for jittering outcomes
  dataset2 <- dataset %>% 
    group_by(get(outcome_var_t)) %>% 
    add_tally() %>% 
    ungroup() %>% 
    mutate(
      # Jitter event times
      jitter = runif(nrow(dataset), min = -.01, max = .01), # N in the samp
      time = ifelse(n>1, get(outcome_var_t) + jitter, get(outcome_var_t))
      )
  
  # Run the AJ model
  aj <- survfit(Surv(time, factor(get(outcome_var))) ~ trt, data = dataset2)
  
  mod <- summary(aj)

  summod <- data.frame(t = mod$time,
                       r = mod$pstate[,2], 
                       #se = mod$std.err[,2],
                       # This is not right yet
                       #trt = c(rep(0, mod$strata["trt=0"]), rep(1, mod$strata["trt=1"]))
                       trt = c(rep(0, length(mod[["strata"]][mod[["strata"]] == "trt=0"])), 
                                   rep(1, length(mod[["strata"]][mod[["strata"]] == "trt=1"])))
                       ) %>%
    filter(t < t_val+0.5) %>%  # Deal with the jittering of outcomes
    group_by(trt) %>% 
    summarize(r = last(r),
              #se = last(se),
              .groups = 'drop') %>% 
    pivot_wider(names_from = trt,
                values_from = c(r), #, se
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           Estimator = "Aalen-Johanssen") %>%
    select(r_0, r_1, rr, #RR_lower, RR_upper, 
           rd #, 
           #RD_lower, RD_upper
           )
  
  return(summod)
    
  
}

## Create a function that calculates the RD and RR that we would have
## -- expected without censoring. This should provide the 
## -- correct estimate for the trial and give a sense of how 
## -- informative the censoring process is.

no_censor_estimator <- function(dataset, outcome_var){
  
  summary <- dataset %>% 
    group_by(trt) %>% 
    summarize(r = sum(get(outcome_var) == 1) / n(),
              .groups = "drop") %>% 
    pivot_wider(names_from = trt,
                values_from = r,
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1 / r_0,
           rd = r_1 - r_0)
  
  return(summary)
  
}



```

## Clean data and conduct analyses

```{r}

trial <- dataset %>% 
  bind_rows() %>% 
  nest(data = c(-sim_id)) %>% 
  mutate(trial_cohort = purrr::map(data, ~trial_cohort(.x)),
         trial_n = purrr::map(trial_cohort, ~total_trial(.x)),
         trial_n_trt = purrr::map(trial_cohort, ~total_trial_trt(.x)),
         trial_indices = purrr::map(trial_cohort, ~trial_index(.x)),
         trial_outcomes = purrr::map(trial_cohort, ~count_outcomes(.x)),
         trial_outcomes_dist = purrr::map(trial_cohort, ~outcomes_dist(.x)),
         trial_pre_sga = purrr::map(trial_cohort, ~count_p_s(.x)),
         
         ## Calculate RRs and risks for composite outcome
         km_composite = purrr::map(trial_cohort, ~km_estimator(.x, outcome_var = "composite_km",
                                                               outcome_var_t = "time", t_val = 41) %>% 
                                     rename_all(~ paste("km_composite_", .))),
         aj_composite = purrr::map(trial_cohort, ~aj_estimator(.x, "composite_multi", 
                                                               "composite_multi_gw", 41) %>% 
                                     rename_all(~ paste("aj_composite_", .))),
         no_censor_composite = purrr::map(trial_cohort, ~no_censor_estimator(.x, outcome_var = "composite_no_censor") %>% 
                                            rename_all(~ paste("no_censor_composite_", .))),
         
         ## Calculate RRs and risks for SGA outcome
         km_sga = purrr::map(trial_cohort, ~km_estimator(.x, outcome_var = "sga_km",
                                                               outcome_var_t = "time", t_val = 41) %>% 
                                     rename_all(~ paste("km_sga_", .))),
         aj_sga_fd = purrr::map(trial_cohort, ~aj_estimator(.x, "sga_multi_do", 
                                                               "sga_multi_do_gw", 41) %>% 
                                     rename_all(~ paste("aj_sga_fd_", .))),
         aj_sga_all = purrr::map(trial_cohort, ~aj_estimator(.x, "sga_multi_all", 
                                                               "sga_multi_all_gw", 41) %>% 
                                     rename_all(~ paste("aj_sga_all_", .))),
         no_censor_sga = purrr::map(trial_cohort, ~no_censor_estimator(.x, outcome_var = "sga_no_censor") %>% 
                                            rename_all(~ paste("no_censor_sga_", .))),
         
         ## Calculate RRs and risks for preterm live birth as outcome
         km_preterm = purrr::map(trial_cohort, ~km_estimator(.x, outcome_var = "preterm_km",
                                                               outcome_var_t = "preterm_t", t_val = 34) %>% 
                                     rename_all(~ paste("km_preterm_", .))),
         aj_preterm = purrr::map(trial_cohort, ~aj_estimator(.x, "preterm_multi", 
                                                               "preterm_multi_gw", 41) %>% 
                                     rename_all(~ paste("aj_preterm_", .))),
         no_censor_preterm = purrr::map(trial_cohort, ~no_censor_estimator(.x, outcome_var = "preterm_no_censor") %>% 
                                          rename_all(~ paste("no_censor_preterm_", .)))
         
         ) %>% 
  unnest_wider(c(km_composite, aj_composite, no_censor_composite, km_sga, aj_sga_fd, aj_sga_all, no_censor_sga,
                 km_preterm, aj_preterm, no_censor_preterm))

## Everything seems to work as expected, though should compare results from original to new

```


# Conduct Analyses

We want to understand the difference in the risk, risk ratio, and risk difference estimates from each analytic approach. This will include the following calculations:

- Risk(Outcome|Treated) via KM - Risk(Outcome|Treated) without Censoring
- Risk(Outcome|Untreated) via KM - Risk(Outcome|Untreated) without Censoring
- Risk(Outcome|Treated) via AJ - Risk(Outcome|Treated) without Censoring
- Risk(Outcome|Untreated) via AJ - Risk(Outcome|Untreated) without Censoring
- Log(RR) via KM - Log(RR) without Censoring
- Log(RR) via Aj - Log(RR) without Censoring
- RD via KM - RD without Censoring
- RD via AJ - RD without Censoring

```{r}

bias_outcome <- function(r1_km, r0_km, r1_nocen, r0_nocen, r1_aj, r0_aj,
                         rr_km, rr_nocen, rr_aj,
                         rd_km, rd_nocen, rd_aj, prefix = ""){
  
  dataset <- tibble(
    risk1_km_nocen = r1_km - r1_nocen,
    risk0_km_nocen = r0_km - r0_nocen,
    risk1_aj_nocen = r1_aj - r1_nocen,
    risk0_aj_nocen = r0_aj - r0_nocen,
    rr_km_nocen = log(rr_km) - log(rr_nocen),
    rr_aj_nocen = log(rr_aj) - log(rr_nocen),
    rd_km_nocen = rd_km - rd_nocen,
    rd_aj_nocen = rd_aj - rd_nocen
  )
  
  dataset <- dataset %>% 
    rename_all(~paste0(prefix, .))
  
  return(dataset)
  
}



trial2 <- trial %>% 
  rowwise() %>% 
  mutate(
    composite = list(bias_outcome(`km_composite_ r_1`, `km_composite_ r_0`, `no_censor_composite_ r_1`,
                             `no_censor_composite_ r_0`, `aj_composite_ r_1`, `aj_composite_ r_0`,
                             `km_composite_ rr`, `no_censor_composite_ rr`, `aj_composite_ rr`,
                             `km_composite_ rd`, `no_censor_composite_ rd`, `aj_composite_ rd`, "composite_")),
    sga_fd = list(bias_outcome(`km_sga_ r_1`, `km_sga_ r_0`, `no_censor_sga_ r_1`, `no_censor_sga_ r_0`,
                               `aj_sga_fd_ r_1`, `aj_sga_fd_ r_0`, 
                               `km_sga_ rr`, `no_censor_sga_ rr`, `aj_sga_fd_ rr`,
                               `km_sga_ rd`, `no_censor_sga_ rd`, `aj_sga_fd_ rd`, "sga_fd_")),
    sga_all = list(bias_outcome(`km_sga_ r_1`, `km_sga_ r_0`, `no_censor_sga_ r_1`, `no_censor_sga_ r_0`,
                               `aj_sga_all_ r_1`, `aj_sga_all_ r_0`, 
                               `km_sga_ rr`, `no_censor_sga_ rr`, `aj_sga_all_ rr`,
                               `km_sga_ rd`, `no_censor_sga_ rd`, `aj_sga_all_ rd`, "sga_all_")),
    preterm = list(bias_outcome(`km_preterm_ r_1`, `km_preterm_ r_0`, `no_censor_preterm_ r_1`,
                                `no_censor_preterm_ r_0`, `aj_preterm_ r_1`, `aj_preterm_ r_0`,
                                `km_preterm_ rr`, `no_censor_preterm_ rr`, `aj_preterm_ rr`,
                                `km_preterm_ rd`, `no_censor_preterm_ rd`, `aj_preterm_ rd`, "preterm_"))
  ) %>% 
  unnest(c(composite, sga_fd, sga_all, preterm))

```

## Summarize Bias

Code for function that summarizes the bias between KM and AJ estimators.

```{r}

compare <- function(km, aj){
  
  # Calculate the differences
  differences <- tibble(
      # Absolute Risks
      r_0 = km$r_0 - aj$r_0,
      r_1 = km$r_1 - aj$r_1,
      # Risk Ratios
      rr = log(km$rr) - log(aj$rr),
      RR_lower = log(km$RR_lower) - log(aj$RR_lower),
      RR_upper = log(km$RR_upper) - log(aj$RR_upper),
      # Risk differences
      rd = km$rd - aj_rrs_composite$rd,
      RD_lower = km$RD_lower - aj$RD_lower,
      RD_upper = km$RD_upper - aj$RD_upper,
      Estimator = "Differences: KM-AJ"
  )
  
  # Make the kable object
  kable_object <- rbind(km, aj) %>% 
    rbind(differences) %>% 
    ungroup() %>% 
    mutate(r_0 = round(r_0, 2),
           r_1 = round(r_1, 2),
           rr = round(rr, 2),
           RR_lower = round(RR_lower, 2),
           RR_upper = round(RR_upper, 2),
           rd = round(rd, 2),
           RD_lower = round(RD_lower, 2),
           RD_upper = round(RD_upper, 2)) %>% 
    #kbl() %>% 
    kable(col.names = c("Estimator", "Risk Among Untreated", "Risk Among Treated","RR", "RR 95% LCL",
                        "RR 95% UCL", "RD", "RD 95% LCL", "RD 95% UCL"),
          align = "lcccccccc") %>% 
    kable_classic() %>% 
    row_spec(3, bold = T, color = "white", background = "black")
  
  return(kable_object)
  
}


```
