---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Identify the cohort that entered the trial.

```{r}


trial <- dataset %>% 
  subset(trial_participant == TRUE)

```

## Number of trial participants, across treatment groups

```{r}

# Approx half and half, as expected
trial %>% 
  group_by(trt) %>% 
  summarize(count = n())

```

## Number of live births and fetal deaths across treatment groups

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
            fd_percent = round(fetal_death / n() * 100,1),
            livebirth = sum(preg_outcome_final == 'livebirth'),
            lb_percent = round(livebirth / n() * 100,1))

```

## Distribution of pregnancy outcomes

Quantitative description of the distribution of pregnancy outcomes, by treatment group.

```{r}

trial %>% 
  group_by(trt, preg_outcome_final) %>% 
  summarize(
            min = min(preg_outcome_final_gw),
            p25 = quantile(preg_outcome_final_gw, 0.25),
            median = median(preg_outcome_final_gw),
            p75 = quantile (preg_outcome_final_gw, 0.75),
            max = max(preg_outcome_final_gw)
            )

# Not sure why we don't see fetal deaths after 34 weeks. Potentially a low prob issue (i.e., not enough n to see those?) - That is where the probability drops substantially in the simulation parameters.

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  ggplot(aes(x = preg_outcome_final_gw, color = preg_outcome_final, fill = preg_outcome_final)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_grid(facets = trt ~.)

```

## Number with preeclampsia and SGA

By treatment group.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(preeclampsia = sum(preeclampsia == 1),
            preec_perc = round(preeclampsia / n() * 100, 1),
            sga = sum(sga == 1),
            sga_perc = round(sga / n() * 100, 1))

```

# Conduct Analyses