---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(Epi)
library(kableExtra)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r, collapse=T}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Write functions that describes the data

```{r}

## Identify the cohort that entered the trial

trial_cohort <- function(dataset){
  
  trial <- dataset %>% 
    subset(trial_participant == TRUE) %>% 
    # Identify the observed outcomes
    mutate(preg_outcome_final = ifelse(trt == 1,
                                       preg_outcome_final_trt,
                                       preg_outcome_final_untrt),
           preg_outcome_final_gw = ifelse(trt == 1,
                                          preg_outcome_final_gw_trt,
                                          preg_outcome_final_gw_untrt),
           preeclampsia = ifelse(trt == 1,
                                 preeclampsia_trt,
                                 preeclampsia_untrt),
           sga = ifelse(trt == 1,
                        sga_trt,
                        sga_untrt)) %>% 
    rowwise() %>% 
    mutate(censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                      1,
                      0),
           time = ifelse(censor == 1,
                         first_censoring_gw,
                         preg_outcome_final_gw)
           ) %>% 
    ungroup() %>% 
    mutate(
      composite_km = case_when(censor == 1 ~ 0,
                          preg_outcome_final == "fetaldeath" & censor == 0 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 1 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 0 ~ 0),
      composite_multi = case_when(censor == 1 ~ 0,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 1,
                                censor == 0 & preeclampsia == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 2
                                ),
      # Make a gestational week variable for the composite_multi variable
      composite_multi_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # Make a SGA variable for the KM estimator
      sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0),
      # Make SGA variables with fetal death only as a competing event
      sga_multi_do = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 0), 
      # Make a gestational week variable for the sga_multi variable
      sga_multi_do_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # Make SGA variables with all competing events
      sga_multi_all = case_when(censor == 1 ~ 0,
                                censor == 0 & sga == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                                censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 2), 
      # COULD BE 3 - COME BACK
      # Make a gestational week variable for the sga_multi variable
      sga_multi_all_gw = ifelse(censor == 1,
                                  first_censoring_gw,
                                  preg_outcome_final_gw),
      # PReterm KM variables
      preterm_km = ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
                          preg_outcome_final == "livebirth", # Previously 37
                          1,
                          0),
      preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
                         ifelse(preterm_km == 0 & time >= 35, 35,
                                time)),
      # Preterm AJ variables - NEED TO CHECK
      preterm_multi = case_when(time >= 35 ~ 0,
                                censor == 1 & time < 35 ~ 0,
                                censor == 0 & preg_outcome_final == "livebirth" & time < 35 ~ 1,
                                censor == 0 & preg_outcome_final == "fetaldeath" & time < 35 ~ 2),
      preterm_multi_gw = ifelse(time >= 35, 35, time)
      # preterm_multi = case_when(censor == 1 ~ 0,
      #                           censor == 0 & preg_outcome_final == "livebirth" & 
      #                             preg_outcome_final_gw < 35 ~ 1,
      #                           censor == 0 & preg_outcome_final == "livebirth" &
      #                             preg_outcome_final_gw >= 35 ~ 2,
      #                           censor == 0 & preg_outcome_final == "fetaldeath" ~ 2),
      # preterm_multi_gw = ifelse(censor == 1,
      #                           first_censoring_gw,
      #                           preg_outcome_final_gw),

      
    )
  
  return(trial)
  
}

    


### Calculate the number of trial participants

total_trial <- function(dataset){
  
  tib <- dataset %>% 
    summarize(count = n()) 

  return(as.double(tib[1,1]))
  
}

## Count the number of patients across treatment groups
## -- Will save this as a tibble

total_trial_trt <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(count = n(),
              percent = round(count / nrow(dataset) * 100),2)
  
  return(tib)
  
}

## Describe indexing into the cohort

trial_index <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(
              min = min(first_index),
              p25 = quantile(first_index, 0.25),
              median = median(first_index),
              p75 = quantile (first_index, 0.75),
              max = max(first_index),
              nmiss = sum(is.na(first_index))
              )
  
  return(tib)
  
}

## Number of live births and fetal deaths across treatment groups
## -- Currently, this does not consider censoring.
count_outcomes <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
              fd_percent = round(fetal_death / n() * 100,1),
              livebirth = sum(preg_outcome_final == 'livebirth'),
              lb_percent = round(livebirth / n() * 100,1))
  
  return(tib)
  
}

## Distribution of pregnancy outcomes
## -- Currently, this does not consider censoring.
outcomes_dist <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt, preg_outcome_final) %>% 
    summarize(
              min = min(preg_outcome_final_gw),
              p25 = quantile(preg_outcome_final_gw, 0.25),
              median = median(preg_outcome_final_gw),
              p75 = quantile (preg_outcome_final_gw, 0.75),
              max = max(preg_outcome_final_gw),
              .groups = 'keep'
              )
  
  return(tib)
  
}

## Number with preeclampsia and SGA
count_p_s <- function(dataset){
  
  tib <- dataset %>% 
    group_by(trt) %>% 
    summarize(preeclampsia = sum(preeclampsia == 1),
              preec_perc = round(preeclampsia / n() * 100, 1),
              sga = sum(sga == 1),
              sga_perc = round(sga / n() * 100, 1))
  
  return(tib)
  
}

## KM Estimator
## -- Code for function that uses the Kaplan-Meier estimator to estimate risk.

km_estimator <- function(dataset, outcome_var, outcome_var_t, t_val){
  
  ## Run the KM analysis only considering the composite outcome.
  km <- survfit(Surv(get(outcome_var_t), get(outcome_var)) ~ trt, data = dataset)
  
  wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, #se = km$std.err,
                          trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))
  
  estimate <- wrisk %>% 
    filter(t < t_val+0.5) %>% # Deal with no additional risk scenarios
    group_by(trt) %>% 
    summarize(r = last(r),
              #se = last(se),
              .groups = 'drop') %>% 
    pivot_wider(names_from = trt,
                values_from = c(r), #, se
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           Estimator = "Kaplan-Meier") %>%
    select(r_0, r_1, rr, #RR_lower, RR_upper, 
           rd #, 
           #RD_lower, RD_upper
           )

  return(estimate)
  
}


## AJ Estimator

aj_estimator <- function(dataset, outcome_var, outcome_var_t, t_val){
  
  #jitter ties
  #set.seed(1234)
  # Not sure if should set seed for jittering outcomes
  dataset2 <- dataset %>% 
    group_by(get(outcome_var_t)) %>% 
    add_tally() %>% 
    ungroup() %>% 
    mutate(
      # Jitter event times
      jitter = runif(nrow(dataset), min = -.01, max = .01), # N in the samp
      time = ifelse(n>1, get(outcome_var_t) + jitter, get(outcome_var_t))
      )
  
  # Run the AJ model
  aj <- survfit(Surv(time, factor(get(outcome_var))) ~ trt, data = dataset2)
  
  mod <- summary(aj)

  summod <- data.frame(t = mod$time,
                       r = mod$pstate[,2], 
                       #se = mod$std.err[,2],
                       # This is not right yet
                       #trt = c(rep(0, mod$strata["trt=0"]), rep(1, mod$strata["trt=1"]))
                       trt = c(rep(0, length(mod[["strata"]][mod[["strata"]] == "trt=0"])), 
                                   rep(1, length(mod[["strata"]][mod[["strata"]] == "trt=1"])))
                       ) %>%
    filter(t < t_val+0.5) %>%  # Deal with the jittering of outcomes
    group_by(trt) %>% 
    summarize(r = last(r),
              #se = last(se),
              .groups = 'drop') %>% 
    pivot_wider(names_from = trt,
                values_from = c(r), #, se
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           Estimator = "Aalen-Johanssen") %>%
    select(r_0, r_1, rr, #RR_lower, RR_upper, 
           rd #, 
           #RD_lower, RD_upper
           )
  
  return(summod)
    
  
}



```

## Clean up the data

```{r}

trial <- dataset %>% 
  bind_rows() %>% 
  nest(data = c(-sim_id)) %>% 
  mutate(trial_cohort = purrr::map(data, ~trial_cohort(.x)),
         trial_n = purrr::map(trial_cohort, ~total_trial(.x)),
         trial_n_trt = purrr::map(trial_cohort, ~total_trial_trt(.x)),
         trial_indices = purrr::map(trial_cohort, ~trial_index(.x)),
         trial_outcomes = purrr::map(trial_cohort, ~count_outcomes(.x)),
         trial_outcomes_dist = purrr::map(trial_cohort, ~outcomes_dist(.x)),
         trial_pre_sga = purrr::map(trial_cohort, ~count_p_s(.x)),
         km_composite = purrr::map(trial_cohort, ~km_estimator(.x, outcome_var = "composite_km",
                                                               outcome_var_t = "time", t_val = 41) %>% 
                                     rename_all(~ paste("km_composite", .))),
         #km_composite = purrr::map()
         aj_composite = purrr::map(trial_cohort, ~aj_estimator(.x, "composite_multi", 
                                                               "composite_multi_gw", 41) %>% 
                                     rename_all(~ paste("aj_composite", .)))) %>% 
  unnest_wider(c(km_composite, aj_composite))

## Everything seems to work as expected, though should compare results from original to new

```


# Conduct Analyses


### AJ Estimator

Code for function that generates the AJ estimator object.

```{r}

aj_estimator <- function(dataset, outcome_var, outcome_var_t){
  
  #jitter ties
  #set.seed(1234)
  # Not sure if should set seed for jittering outcomes
  dataset2 <- dataset %>% 
    group_by(get(outcome_var_t)) %>% 
    add_tally() %>% 
    ungroup() %>% 
    mutate(
      # Jitter event times
      jitter = runif(nrow(trial2), min = -.01, max = .01), # N in the samp
      time = ifelse(n>1, get(outcome_var_t) + jitter, get(outcome_var_t))
      )
  
  # Run the AJ model
  aj <- survfit(Surv(time, factor(get(outcome_var))) ~ trt, data = dataset2)
  
}

#test <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

```

Code for function that generates RRs from the AJ estimator.

```{r}
#ASK JESS ABOUT t_val specification

aj_rr <- function(aj_object, t_val){
  
  mod <- summary(aj_object)

  summod <- data.frame(t = mod$time,
                       r = mod$pstate[,2], 
                       se = mod$std.err[,2],
                       # This is not right yet
                       #trt = c(rep(0, mod$strata["trt=0"]), rep(1, mod$strata["trt=1"]))
                       trt = c(rep(0, length(mod[["strata"]][mod[["strata"]] == "trt=0"])), 
                                   rep(1, length(mod[["strata"]][mod[["strata"]] == "trt=1"])))
                       ) %>%
    filter(t < t_val+0.5) %>%  # Deal with the jittering of outcomes
    group_by(trt) %>% 
    summarize(r = last(r),
              se = last(se),
              .groups = 'drop') %>% 
    select(r,se,trt) %>% 
      pivot_wider(names_from = trt,
                  values_from = c(r, se),
                  names_glue = "{.value}_{trt}") %>% 
      rowwise() %>% 
      mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           # use delta method to get 95% CI
           logRR = log(rr),
           var_1 = se_1^2,
           var_0 = se_0^2,
           var_logRR = (1/r_0)^2 * var_0 + (1/r_1)^2 * var_1,
           logRR_lower = logRR - 1.96*sqrt(var_logRR),
           logRR_upper = logRR + 1.96*sqrt(var_logRR),
           RR_lower = exp(logRR_lower),
           RR_upper = exp(logRR_upper),
           se_rd = sqrt(se_1^2 + se_0^2),
           RD_lower = rd - 1.96*se_rd,
           RD_upper = rd + 1.96*se_rd,
           Estimator = "Aalen-Johansen") %>%
    select(Estimator, r_0, r_1, rr, RR_lower, RR_upper, rd, RD_lower, RD_upper)
  
  return(summod)
    
}

# aj_rr(test)

```

Code for function that plots the AJ object. **This still needs some work to be sure that it's outputting everything correctly.** Currently, the x-axis is wrong. Should be from index date.

```{r}

aj_plot <- function(aj_object, xlab="Gestational weeks post-conception", ylab="Probability in State",
                    outcome = "", competing = ""){
  plot(aj_object, col=c(1,2,1,2), lty=c(2,2,1,1),
       mark.time=FALSE, lwd=2,  #xscale=12,
       xlab=xlab, ylab=ylab)
  legend("topleft", legend = c(paste(competing, ":untreated"), paste(competing, ":treated"),
                               paste(outcome, ":untreated"), paste(outcome, ":treated") 
                               ),
         col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

  # legend(240, .6, c(paste(outcome, ":untreated"), paste(outcome, ":treated"), paste(competing, ":untreated"), paste(competing, ":treated")),
  #        col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

}

#aj_plot(test, outcome = "Composite", competing = "Delivery")


## VERY HELPFUL: https://cran.r-project.org/web/packages/survival/vignettes/compete.pdf

## Plotting AJ estimators:
#https://search.r-project.org/CRAN/refmans/Epi/html/plotCIF.html

## More AJ R code:
#https://cran.r-project.org/web/packages/AalenJohansen/index.html

```

## Summarize Bias

Code for function that summarizes the bias between KM and AJ estimators.

```{r}

compare <- function(km, aj){
  
  # Calculate the differences
  differences <- tibble(
      # Absolute Risks
      r_0 = km$r_0 - aj$r_0,
      r_1 = km$r_1 - aj$r_1,
      # Risk Ratios
      rr = log(km$rr) - log(aj$rr),
      RR_lower = log(km$RR_lower) - log(aj$RR_lower),
      RR_upper = log(km$RR_upper) - log(aj$RR_upper),
      # Risk differences
      rd = km$rd - aj_rrs_composite$rd,
      RD_lower = km$RD_lower - aj$RD_lower,
      RD_upper = km$RD_upper - aj$RD_upper,
      Estimator = "Differences: KM-AJ"
  )
  
  # Make the kable object
  kable_object <- rbind(km, aj) %>% 
    rbind(differences) %>% 
    ungroup() %>% 
    mutate(r_0 = round(r_0, 2),
           r_1 = round(r_1, 2),
           rr = round(rr, 2),
           RR_lower = round(RR_lower, 2),
           RR_upper = round(RR_upper, 2),
           rd = round(rd, 2),
           RD_lower = round(RD_lower, 2),
           RD_upper = round(RD_upper, 2)) %>% 
    #kbl() %>% 
    kable(col.names = c("Estimator", "Risk Among Untreated", "Risk Among Treated","RR", "RR 95% LCL",
                        "RR 95% UCL", "RD", "RD 95% LCL", "RD 95% UCL"),
          align = "lcccccccc") %>% 
    kable_classic() %>% 
    row_spec(3, bold = T, color = "white", background = "black")
  
  return(kable_object)
  
}


```

## Composite Preeclampsia and Fetal Death

### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}
# 
# ## Create the composite outcome
# trial2 <- trial %>% 
#   mutate(
#     # Make an indicator that the patient was censored before we observed their outcome - made earlier
#     # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
#     #                 1,
#     #                 0),
#     # Make composite
#     composite = case_when(censor == 1 ~ 0,
#                           preg_outcome_final == "fetaldeath" & censor == 0 ~ 1,
#                           preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 1 ~ 1,
#                           preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 0 ~ 0) #,
#     # composite = ifelse(preg_outcome_final == "fetaldeath" | preeclampsia == 1 & censor == 0,
#     #                         1,
#     #                         0),
#     # composite_gw = ifelse(censor == 0,
#     #                       preg_outcome_final_gw,
#     #                       first_censoring_gw)
#     )

## Run the KM analysis only considering the composite outcome.
#wrisk <- km_estimator(trial2, "composite","composite_gw")
wrisk <- km_estimator(trial2, "composite","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for Composite Outcome")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, heigh = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

#############################################################################################
## Calculate the RR at different weeks of follow-up; SE is infinity at 42 weeks of follow-up
## -- ASK JESS about this.
## -- Funky thing: The standard errors get substantially larger as the weeks progress.
## -- Likely because there are fewer and fewer individuals at risk at 42 weeks and imputing
## -- more censoring.
## -- Need to check this intuition with Jess.

km_rr <- km_rr_wCI(wrisk, 41)


```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make a multi-level outcome
    composite_multi = case_when(censor == 1 ~ 0,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 1,
                                censor == 0 & preeclampsia == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 2
                                ),
    # Make a gestational week variable for the composite_multi variable
    composite_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "composite", competing = "delivery")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate risks without censoring

```{r}

trial3 <- trial %>% 
  mutate(composite = case_when(preg_outcome_final == "fetaldeath" ~ 1,
                               preg_outcome_final == "livebirth" & preeclampsia == 1 ~ 1,
                               preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(composite == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## SGA with Fetal Death Only as Competing Event


### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make an indicator for sga, incorporating censoring
    sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0)#,
    # Make a time variable for sga, incorporating censoring
    # sga_t = ifelse(sga == 1 & censor == 0,
    #                preg_outcome_final_gw,
    #                first_censoring_gw)
    )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "sga_km","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for SGA")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

# Undefined at 42 weeks
km_rr <- km_rr_wCI(wrisk, 42)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    sga_multi = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 0), 
    # COULD BE 3 - COME BACK
    # Make a gestational week variable for the sga_multi variable
    sga_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "sga_multi", "sga_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "sga", competing = "no sga")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(sga_out = ifelse(sga == 1, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(sga_out == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## SGA with all Competing Events

### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make an indicator for sga, incorporating censoring
    sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0)#,
    # Make a time variable for sga, incorporating censoring
    # sga_t = ifelse(sga == 1 & censor == 0,
    #                preg_outcome_final_gw,
    #                first_censoring_gw)
    )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "sga_km","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for SGA")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

# Undefined at 42 weeks
km_rr <- km_rr_wCI(wrisk, 42)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    sga_multi = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 2), 
    # COULD BE 3 - COME BACK
    # Make a gestational week variable for the sga_multi variable
    sga_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "sga_multi", "sga_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "sga", competing = "no sga")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(sga_out = ifelse(sga == 1, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(sga_out == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## Preterm live birth

### KM Estimator

Output the cumulative incidence curves at the end of follow-up. Note that we generated gestational weeks from conception, not LMP. Preterm birth is defined as birth <37 weeks of gestation from LMP (2 weeks prior to conception). Thus, we define preterm birth here as birth that occurs at <35 weeks of gestation from conception.

```{r}

## Create the preterm live birth outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & 
                      !is.na(first_censoring_gw),
                    1,
                    0),
    # Make an indicator for sga, incorporating censoring
    preterm_km = ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
                          preg_outcome_final == "livebirth", # Previously 37
                        1,
                        0),
    preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
                       ifelse(preterm_km == 0 & time >= 35, 34,
                              time))
    # preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
    #                ifelse(preg_outcome_final_gw < 35 & preg_outcome_final == "fetaldeath" & 
    #                         censor == 0, preg_outcome_final_gw,
    #                       ifelse(preg_outcome_final_gw >= 35 & censor == 0, 34,
    #                              ifelse(censor == 1 & first_censoring_gw >= 35, 34,
    #                                     ifelse(censor == 1 & first_censoring_gw < 35, 
    #                                            first_censoring_gw, NA)))))

    # preterm_t = case_when(
    #   preterm_km == 1 ~ preg_outcome_final_gw,
    #   preg_outcome_final_gw < 35 & preg_outcome_final == "fetaldeath" & censor == 0 ~ preg_outcome_final_gw,
    #   preg_outcome_final_gw >= 35 & censor == 0 ~ 35,
    #   censor == 1 & first_censoring_gw >= 35 ~ 35,
    #   censor == 1 & first_censoring_gw < 35 ~ first_censoring_gw
    # )

  )


      
  #     ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
  #                        preg_outcome_final == "livebirth", # Previously 37
  #                      preg_outcome_final_gw,
  #                      first_censoring_gw)
  #   ) %>% 
  # # Now, account for the fact that follow-up should stop at 36 weeks.
  # mutate(
  #   # Reset outcome to what it should be at 36 weeks
  #   # This functionally shouldn't matter since preterm can only happen before 36 weeks
  #   preterm_km = ifelse(preterm_t >= 35,
  #                       0,
  #                       preterm_km),
  #   # Reset timing so that it ends at 36 weeks of follow-up
  #   preterm_t = ifelse(preterm_t >= 35,
  #                      35,
  #                      preterm_t)
  # )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "preterm_km","preterm_t")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for Preterm Live Birth")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

km_rr <- km_rr_wCI(wrisk, 36)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    preterm_multi = case_when(censor == 1 ~ 0,
                              censor == 0 & preg_outcome_final == "livebirth" & 
                                preg_outcome_final_gw < 35 ~ 1,
                              censor == 0 & preg_outcome_final == "livebirth" &
                                preg_outcome_final_gw >= 35 ~ 2,
                              censor == 0 & preg_outcome_final == "fetaldeath" ~ 2),
    # Don't think we would calculate risks after 36 weeks since no longer at risk. Fetal 
    # -- death is only competing event.
    # However,  do need to leave them in if not limiting all of the data to 36 weeks. 
    # -- Need to clean up these data more.
    # Make a gestational week variable for the sga_multi variable
    preterm_multi_gw = ifelse(censor == 1,
                              first_censoring_gw,
                              preg_outcome_final_gw)
  ) %>% 
  #Now, account for the fact that follow-up should stop at 36 weeks
  mutate(
    # Fix the outcomes to no event (i.e., censor) if the event timing is after 36 weeks
    preterm_multi = ifelse(preterm_multi_gw >= 35,
                           0, # i.e., treat as censoring event.
                          # 2, #i.e., safe from preterm birth, even if ultimatley end in fetal death
                           preterm_multi),
    # Cap the event timing at 36 weeks
    preterm_multi_gw = ifelse(preterm_multi_gw >= 35,
                              36, # i.e., if treat as censoring event
                              #36,
                              preterm_multi_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "preterm_multi", "preterm_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 36)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "preterm", competing = "fetal death")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(preterm = ifelse(preg_outcome_final == "livebirth" & preg_outcome_final_gw < 35, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(preterm == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```