---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Identify the cohort that entered the trial.

```{r}


trial <- dataset %>% 
  subset(trial_participant == TRUE) %>% 
  # Identify the observed outcomes
  mutate(preg_outcome_final = ifelse(trt == 1,
                                     preg_outcome_final_trt,
                                     preg_outcome_final_untrt),
         preg_outcome_final_gw = ifelse(trt == 1,
                                        preg_outcome_final_gw_trt,
                                        preg_outcome_final_gw_untrt),
         preeclampsia = ifelse(trt == 1,
                               preeclampsia_trt,
                               preeclampsia_untrt),
         sga = ifelse(trt == 1,
                      sga_trt,
                      sga_untrt))

```

## Number of trial participants

### Total

```{r}

trial %>% 
  summarize(count = n())

```

### Across treatment groups

```{r}

# Approx half and half, as expected
trial %>% 
  group_by(trt) %>% 
  summarize(count = n(),
            percent = round(count / nrow(trial) * 100),2)

```

### Timing of indexing into the cohort

Indexing was not intentionally differential by treatment group. If differences occur, it is likely due to patterns of pregnancy loss. This may not be particularly dramatic in the setting where everyone has a 50/50 chance of indexing every gestational week from weeks 4 through 20 (from conception).

Quantitative description of the distribution of gestational age at trial indices, by treatment group.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(
            min = min(first_index),
            p25 = quantile(first_index, 0.25),
            median = median(first_index),
            p75 = quantile (first_index, 0.75),
            max = max(first_index)
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  ggplot(aes(x = first_index)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_wrap(~trt_label, scales = "free") +
  ylim(0, 200) +  # Set the y-axis limit to 0 to 200
  xlim(4, 20) +
  xlab("Gestational Week at Index") +  # Set the x-axis label
  ylab("Count")

```

## Number of live births and fetal deaths across treatment groups

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
            fd_percent = round(fetal_death / n() * 100,1),
            livebirth = sum(preg_outcome_final == 'livebirth'),
            lb_percent = round(livebirth / n() * 100,1))

```

## Distribution of pregnancy outcomes

Quantitative description of the distribution of pregnancy outcomes, by treatment group.

```{r}

trial %>% 
  group_by(trt, preg_outcome_final) %>% 
  summarize(
            min = min(preg_outcome_final_gw),
            p25 = quantile(preg_outcome_final_gw, 0.25),
            median = median(preg_outcome_final_gw),
            p75 = quantile (preg_outcome_final_gw, 0.75),
            max = max(preg_outcome_final_gw),
            .groups = 'keep'
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  ggplot(aes(x = preg_outcome_final_gw, color = preg_outcome_final, fill = preg_outcome_final)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_grid(facets = trt_label ~.) +
  xlab("Gestational Week at Pregnancy Outcome") +
  ylab("Count") +
  labs(fill = "Pregnancy Outcome") 

```

## Number with preeclampsia and SGA

By treatment group.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(preeclampsia = sum(preeclampsia == 1),
            preec_perc = round(preeclampsia / n() * 100, 1),
            sga = sum(sga == 1),
            sga_perc = round(sga / n() * 100, 1))

```

# Conduct Analyses

## Composite Preeclampsia and Fetal Death

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(composite = ifelse(preg_outcome_final == "fetaldeath" || preeclampsia == 1,
                            1,
                            0))

## Run the KM analysis only considering the composite outcome.
km <- survfit(Surv(preg_outcome_final_gw, composite) ~ trt, data = trial2)
#summary(km)

wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, 
                        trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))
riskplot <- ggplot() +
  geom_step(data = wrisk, aes(x = t, y = r,  group = factor(trt), color = factor(trt)))+
  xlab("Weeks")+
  ylab("Risk") +
  scale_color_discrete(labels = c("Treated", "Untreated"))+
 # scale_y_continuous(limits = c(0, 0.55), breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5))+
  #scale_x_continuous(limits = c(0, 12), breaks = c(0, 2, 4, 6, 8,10, 12))+
  theme_classic(base_size = 15)+
  theme(legend.position = c(0.28, 0.9), legend.title = element_blank(),
        legend.background = element_rect(fill="transparent"), 
        panel.background = element_rect(fill = NA), 
        panel.grid.major = element_line(colour = "gray", linetype = "dotted"))

riskplot

ggsave(riskplot, file = "q13.png", width = 5, heigh = 5, units = "cm")


#get rr
rr <- wrisk %>% 
  group_by(trt) %>% 
  summarize(risk = last(r)) %>% 
  mutate(rr = risk/lag(risk, default =NA)) %>% 
  filter(rr == last(rr)) %>% select(rr) %>% unlist()
rr


## Run the competing events analysis
## -- We haven't introduced censoring right now, so there are only two possible outcomes:
## -- fetal death or live birth.

#jitter ties
trial2 <- trial2 %>% 
  group_by(preg_outcome_final_gw) %>% 
  add_tally() %>% 
  ungroup() %>% 
  mutate(
    jitter = runif(nrow(trial2), min = -.01, max = .01), # N in the samp
    preg_outcome_final_gw = ifelse(n>1, preg_outcome_final_gw + jitter, preg_outcome_final_gw))



```

## SGA

## Preterm live birth

