---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(Epi)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Identify the cohort that entered the trial.

```{r}


trial <- dataset %>% 
  subset(trial_participant == TRUE) %>% 
  # Identify the observed outcomes
  mutate(preg_outcome_final = ifelse(trt == 1,
                                     preg_outcome_final_trt,
                                     preg_outcome_final_untrt),
         preg_outcome_final_gw = ifelse(trt == 1,
                                        preg_outcome_final_gw_trt,
                                        preg_outcome_final_gw_untrt),
         preeclampsia = ifelse(trt == 1,
                               preeclampsia_trt,
                               preeclampsia_untrt),
         sga = ifelse(trt == 1,
                      sga_trt,
                      sga_untrt))

# Note that censoring is the same regardless of treatment.

```

## Number of trial participants

### Total

```{r}

trial %>% 
  summarize(count = n())

```

### Across treatment groups

```{r}

# Approx half and half, as expected
trial %>% 
  group_by(trt) %>% 
  summarize(count = n(),
            percent = round(count / nrow(trial) * 100),2)

```

### Timing of indexing into the cohort

Indexing was not intentionally differential by treatment group. If differences occur, it is likely due to patterns of pregnancy loss. This may not be particularly dramatic in the setting where everyone has a 50/50 chance of indexing every gestational week from weeks 4 through 20 (from conception).

Quantitative description of the distribution of gestational age at trial indices, by treatment group.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(
            min = min(first_index),
            p25 = quantile(first_index, 0.25),
            median = median(first_index),
            p75 = quantile (first_index, 0.75),
            max = max(first_index)
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  ggplot(aes(x = first_index)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_wrap(~trt_label, scales = "free") +
  ylim(0, 200) +  # Set the y-axis limit to 0 to 200
  xlim(4, 20) +
  xlab("Gestational Week at Index") +  # Set the x-axis label
  ylab("Count")

```

## Number of live births and fetal deaths across treatment groups

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
            fd_percent = round(fetal_death / n() * 100,1),
            livebirth = sum(preg_outcome_final == 'livebirth'),
            lb_percent = round(livebirth / n() * 100,1))

```

## Distribution of pregnancy outcomes

Quantitative description of the distribution of pregnancy outcomes, by treatment group.

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt, preg_outcome_final) %>% 
  summarize(
            min = min(preg_outcome_final_gw),
            p25 = quantile(preg_outcome_final_gw, 0.25),
            median = median(preg_outcome_final_gw),
            p75 = quantile (preg_outcome_final_gw, 0.75),
            max = max(preg_outcome_final_gw),
            .groups = 'keep'
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  ggplot(aes(x = preg_outcome_final_gw, color = preg_outcome_final, fill = preg_outcome_final)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_grid(facets = trt_label ~.) +
  xlab("Gestational Week at Pregnancy Outcome") +
  ylab("Count") +
  labs(fill = "Pregnancy Outcome") 

```

## Number with preeclampsia and SGA

By treatment group.

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(preeclampsia = sum(preeclampsia == 1),
            preec_perc = round(preeclampsia / n() * 100, 1),
            sga = sum(sga == 1),
            sga_perc = round(sga / n() * 100, 1))

```

# Conduct Analyses

## KM Estimator Functions

Write out a function that will use the Kaplan-Meier estimator to estimate risk.

```{r}

km_estimator <- function(dataset, outcome_var, outcome_var_t){
  
  ## Run the KM analysis only considering the composite outcome.
  km <- survfit(Surv(get(outcome_var_t), get(outcome_var)) ~ trt, data = dataset)
  #summary(km)
  
  wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, se = km$std.err,
                          trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))
  
  return(wrisk)
  
}

```

Write out a function that will plot the KM curve.

```{r}

km_plot <- function(dataset, plot_title){
  riskplot <- ggplot() +
    geom_step(data = dataset, aes(x = t, y = r,  group = factor(trt), color = factor(trt)))+
    xlab("Weeks")+
    ylab("Risk") +
    scale_color_discrete(labels = c("Treated", "Untreated"))+
    ggtitle(plot_title) +  # Add the title parameter
   # scale_y_continuous(limits = c(0, 0.55), breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5))+
    #scale_x_continuous(limits = c(0, 12), breaks = c(0, 2, 4, 6, 8,10, 12))+
    theme_classic(base_size = 15)+
    theme(legend.position = c(0.28, 0.9), legend.title = element_blank(),
          legend.background = element_rect(fill="transparent"), 
          panel.background = element_rect(fill = NA), 
          panel.grid.major = element_line(colour = "gray", linetype = "dotted")) 
  
  return(riskplot)
}


```

Write out a function to get the Risk Ratio at different points in follow-up

```{r, collapse=TRUE}


# Function to get the risk ratio with the corresponding confidence interval
km_rr_wCI <- function(data, t_val){
  
  estimate <- data %>% 
    filter(t == t_val) %>% 
    select(r,se,trt) %>% 
    pivot_wider(names_from = trt,
                values_from = c(r, se),
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1 / r_0,
           selnrr = sqrt((1/r_0)^2 * se_0^2 + (1/r_1)^2*se_1^2),
           lcl = exp(log(rr) - 1.96*selnrr),
           ucl = exp(log(rr) + 1.96*selnrr)) %>% 
    select(r_0, r_1, rr, lcl, ucl)
  
  return(estimate)
  
}

#CHECK SE ESTIMATE FOR KM RR

```

## AJ Estimator Functions

Write a function that generates the AJ estimator object.

```{r}

aj_estimator <- function(dataset, outcome_var, outcome_var_t){
  
  #jitter ties
  set.seed(1234) # Set seed so that censoring is the same
  dataset2 <- dataset %>% 
    group_by(get(outcome_var_t)) %>% 
    add_tally() %>% 
    ungroup() %>% 
    mutate(
      # Jitter event times
      jitter = runif(nrow(trial2), min = -.01, max = .01), # N in the samp
      time = ifelse(n>1, get(outcome_var_t) + jitter, get(outcome_var_t))
      )
  
  # Run the AJ model
  aj <- survfit(Surv(time, factor(get(outcome_var))) ~ trt, data = dataset2)
  
}

#test <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

```

Write a function that generates RRs from the AJ estimator.

```{r}

aj_rr <- function(aj_object){
  
  mod <- summary(aj_object)

  summod <- data.frame(t = mod$time,
                       r = mod$pstate[,2], 
                       se = mod$std.err[,2],
                       # This is not right yet
                       #trt = c(rep(0, mod$strata["trt=0"]), rep(1, mod$strata["trt=1"]))
                       trt = c(rep(0, length(mod[["strata"]][mod[["strata"]] == "trt=0"])), 
                                   rep(1, length(mod[["strata"]][mod[["strata"]] == "trt=1"])))
                       ) %>%
    group_by(trt) %>% 
    summarize(r = last(r),
              se = last(se),
              .groups = 'drop') %>% 
    select(r,se,trt) %>% 
      pivot_wider(names_from = trt,
                  values_from = c(r, se),
                  names_glue = "{.value}_{trt}") %>% 
      rowwise() %>% 
      mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           # use delta method to get 95% CI
           logRR = log(rr),
           var_1 = se_1^2,
           var_0 = se_0^2,
           var_logRR = (1/r_0)^2 * var_0 + (1/r_1)^2 * var_1,
           logRR_lower = logRR - 1.96*sqrt(var_logRR),
           logRR_upper = logRR + 1.96*sqrt(var_logRR),
           RR_lower = exp(logRR_lower),
           RR_upper = exp(logRR_upper),
           se_rd = sqrt(se_1^2 + se_0^2),
           RD_lower = rd - 1.96*se_rd,
           RD_upper = rd + 1.96*se_rd) %>%
    select(rr, RR_lower, RR_upper, rd, RD_lower, RD_upper)
  
  return(summod)
    
}

# aj_rr(test)

```

Write a function that plots the AJ object.

This still needs some work to be sure that it's outputting everything correctly.

```{r}

aj_plot <- function(aj_object, xlab="Gestational weeks post-conception", ylab="Probability in State",
                    outcome = "", competing = ""){
  plot(aj_object, col=c(1,2,1,2), lty=c(2,2,1,1),
       mark.time=FALSE, lwd=2,  xscale=12,
       xlab=xlab, ylab=ylab)
    legend("topleft", legend = c(paste(outcome, ":untreated"), paste(outcome, ":treated"), 
                                 paste(competing, ":untreated"), paste(competing, ":treated")),
         col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

  # legend(240, .6, c(paste(outcome, ":untreated"), paste(outcome, ":treated"), paste(competing, ":untreated"), paste(competing, ":treated")),
  #        col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

}

#aj_plot(test, outcome = "Composite", competing = "Delivery")





## VERY HELPFUL: https://cran.r-project.org/web/packages/survival/vignettes/compete.pdf

## Plotting AJ estimators:
#https://search.r-project.org/CRAN/refmans/Epi/html/plotCIF.html

## More AJ R code:
#https://cran.r-project.org/web/packages/AalenJohansen/index.html






```

###

## Composite Preeclampsia and Fetal Death

### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
set.seed(1234) # Set seed so that censoring is the same
trial2 <- trial %>% 
  mutate(composite = ifelse(preg_outcome_final == "fetaldeath" | preeclampsia == 1 & 
                              (is.na(first_censoring_gw) | preg_outcome_final_gw < first_censoring_gw),
                            1,
                            0),
         composite_gw = ifelse(is.na(first_censoring_gw) | preg_outcome_final_gw < first_censoring_gw,
                               preg_outcome_final_gw,
                               first_censoring_gw)
         )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "composite","composite_gw")


# km <- survfit(Surv(preg_outcome_final_gw, composite) ~ trt, data = trial2)
# #summary(km)
# 
# wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, se = km$std.err,
#                         trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))

riskplot <- km_plot(wrisk, "KM Estimator of Risk for Composite Outcome")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, heigh = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

#############################################################################################
## Calculate the RR at different weeks of follow-up; SE is infinity at 42 weeks of follow-up
## -- ASK JESS about this.
## -- Funky thing: The standard errors get substantially larger as the weeks progress.
## -- Likely because there are fewer and fewer individuals at risk at 42 weeks.
## -- Need to check this intuition with Jess.

# Get it at 39 weeks
# km_rr_wCI(wrisk, 39)

# 40 weeks
# km_rr_wCI(wrisk, 40)

# 41 weeks
# km_rr_wCI(wrisk, 41)

# 42 weeks post-LMP, or 40 weeks post-conception
km_rr_wCI(wrisk, 42)


## Hand calculate at 42 weeks
## This is a closed cohort without LTFU, this should be correct
## -- regardless of AJ estimator.
## -- Would be interesting to look at the differences in their risk estimates.
trial2 %>% 
  group_by(trt) %>% 
  summarize(count = sum(composite == 1),
            percent = round(count / n() * 100,2),
            .groups = 'drop') 
log(0.3659/0.4766) - log(0.4268/0.5995)



```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    composite_multi = case_when(censor == 1 ~ 0,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 1,
                                censor == 0 & preeclampsia == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 2
                                ),
    # Make a gestational week variable for the composite_multi variable
    composite_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj)

aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "composite", competing = "delivery")




```

## SGA

## Preterm live birth

