---
title: "Analyses"
author: "Chase Latour"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(Epi)
library(kableExtra)
```

# Description

This HTML/Rmd document will contain the analyses for the data generated for the multiple_outcomes_pregnancy R project.

# Describe Cohort

Describe the patient cohort. This will allow us to check for weird data issues.

## Read in the data

```{r, collapse=T}

# Get the directory where the Rmd file is located
rmd_directory <- dirname(rstudioapi::getSourceEditorContext()$path)

# Specify the Excel file name and path relative to the Rmd file location
data_read <- file.path(rmd_directory, "data.rds")

dataset <- readRDS(data, file = data_read)

```

## Identify the cohort that entered the trial.

```{r}


trial <- dataset %>% 
  subset(trial_participant == TRUE) %>% 
  # Identify the observed outcomes
  mutate(preg_outcome_final = ifelse(trt == 1,
                                     preg_outcome_final_trt,
                                     preg_outcome_final_untrt),
         preg_outcome_final_gw = ifelse(trt == 1,
                                        preg_outcome_final_gw_trt,
                                        preg_outcome_final_gw_untrt),
         preeclampsia = ifelse(trt == 1,
                               preeclampsia_trt,
                               preeclampsia_untrt),
         sga = ifelse(trt == 1,
                      sga_trt,
                      sga_untrt)) %>% 
  rowwise() %>% 
  mutate(censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
         time = ifelse(censor == 1,
                       first_censoring_gw,
                       preg_outcome_final_gw)
         ) %>% 
  ungroup()

# Note that censoring is the same regardless of treatment.


```

## Number of trial participants

### Total

```{r}

trial %>% 
  summarize(count = n())

```

### Across treatment groups

```{r}

# Approx half and half, as expected
trial %>% 
  group_by(trt) %>% 
  summarize(count = n(),
            percent = round(count / nrow(trial) * 100),2)

```

### Timing of indexing into the cohort

Indexing was not intentionally differential by treatment group. If differences occur, it is likely due to patterns of pregnancy loss. This may not be particularly dramatic in the setting where everyone has a 50/50 chance of indexing every gestational week from weeks 4 through 20 (from conception).

Quantitative description of the distribution of gestational age at trial indices, by treatment group.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(
            min = min(first_index),
            p25 = quantile(first_index, 0.25),
            median = median(first_index),
            p75 = quantile (first_index, 0.75),
            max = max(first_index),
            nmiss = sum(is.na(first_index))
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  #mutate(trt_label = factor(trt)) %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  ggplot(aes(x = first_index)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_wrap(~trt_label, scales = "free") +
  #ylim(0, 200) +  # Set the y-axis limit to 0 to 200
  xlim(1, 19) +
  xlab("Gestational Week at Index") +  # Set the x-axis label
  ylab("Count")

```

## Number of live births and fetal deaths across treatment groups

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(fetal_death = sum(preg_outcome_final == 'fetaldeath'),
            fd_percent = round(fetal_death / n() * 100,1),
            livebirth = sum(preg_outcome_final == 'livebirth'),
            lb_percent = round(livebirth / n() * 100,1))

```

## Distribution of pregnancy outcomes

Quantitative description of the distribution of pregnancy outcomes, by treatment group.

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt, preg_outcome_final) %>% 
  summarize(
            min = min(preg_outcome_final_gw),
            p25 = quantile(preg_outcome_final_gw, 0.25),
            median = median(preg_outcome_final_gw),
            p75 = quantile (preg_outcome_final_gw, 0.75),
            max = max(preg_outcome_final_gw),
            .groups = 'keep'
            )

```

Figural description of the distribution of pregnancy outcomes, facet wrapped by treatment group.

```{r}

trial %>% 
  mutate(trt_label = factor(trt, labels = c("Treatment = 0", "Treatment = 1"))) %>% 
  #mutate(trt_label = factor(trt)) %>% 
  ggplot(aes(x = preg_outcome_final_gw, fill = preg_outcome_final)) +
  geom_histogram(alpha=0.35, position = 'identity', binwidth = 1) +
  facet_grid(facets = trt_label ~.) +
  xlab("Gestational Week at Pregnancy Outcome") +
  ylab("Count") +
  labs(fill = "Pregnancy Outcome") 

```

## Number with preeclampsia and SGA

By treatment group.

Currently, these counts ignore censoring events.

```{r}

trial %>% 
  group_by(trt) %>% 
  summarize(preeclampsia = sum(preeclampsia == 1),
            preec_perc = round(preeclampsia / n() * 100, 1),
            sga = sum(sga == 1),
            sga_perc = round(sga / n() * 100, 1))

```

# Conduct Analyses

## Write Functions

### KM Estimator

Code for function that uses the Kaplan-Meier estimator to estimate risk.

```{r}

km_estimator <- function(dataset, outcome_var, outcome_var_t){
  
  ## Run the KM analysis only considering the composite outcome.
  km <- survfit(Surv(get(outcome_var_t), get(outcome_var)) ~ trt, data = dataset)
  # Test: km <- survfit(Surv(preterm_t, preterm_km) ~ trt, data = trial2)
  #summary(km)
  
  wrisk <- data.frame(t = km$time, s = km$surv, r = 1 - km$surv, se = km$std.err,
                          trt = c(rep(0, km$strata["trt=0"]), rep(1, km$strata["trt=1"])))
  
  return(wrisk)
  
}

```

Code for function that plots the KM curve.

```{r}

km_plot <- function(dataset, plot_title){
  riskplot <- ggplot() +
    geom_step(data = dataset, aes(x = t, y = r,  group = factor(trt), color = factor(trt)))+
    xlab("Weeks from Index")+
    ylab("Risk") +
    scale_color_discrete(labels = c("Untreated", "Treated"))+
    ggtitle(plot_title) +  # Add the title parameter
   # scale_y_continuous(limits = c(0, 0.55), breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5))+
    #scale_x_continuous(limits = c(2, 12), breaks = c(0, 2, 4, 6, 8,10, 12))+
    theme_classic(base_size = 15)+
    theme(legend.position = c(0.28, 0.9), legend.title = element_blank(),
          legend.background = element_rect(fill="transparent"), 
          panel.background = element_rect(fill = NA), 
          panel.grid.major = element_line(colour = "gray", linetype = "dotted")) 
  
  return(riskplot)
}


```

Code for function to get the Risk Ratio at different points in follow-up.

```{r, collapse=TRUE}


# Function to get the risk ratio with the corresponding confidence interval
km_rr_wCI <- function(data, t_val){
  
  estimate <- data %>% 
    filter(t < t_val+0.5) %>% # Deal with no additional risk scenarios
    group_by(trt) %>% 
    summarize(r = last(r),
              se = last(se),
              .groups = 'drop') %>% 
    # filter(t == t_val) %>% 
    # select(r, se, trt) %>% 
    pivot_wider(names_from = trt,
                values_from = c(r, se),
                names_glue = "{.value}_{trt}") %>% 
    rowwise() %>% 
    mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           # use delta method to get 95% CI
           logRR = log(rr),
           var_1 = se_1^2,
           var_0 = se_0^2,
           var_logRR = (1/r_0)^2 * var_0 + (1/r_1)^2 * var_1,
           logRR_lower = logRR - 1.96*sqrt(var_logRR),
           logRR_upper = logRR + 1.96*sqrt(var_logRR),
           RR_lower = exp(logRR_lower),
           RR_upper = exp(logRR_upper),
           se_rd = sqrt(se_1^2 + se_0^2),
           RD_lower = rd - 1.96*se_rd,
           RD_upper = rd + 1.96*se_rd,
           Estimator = "Kaplan-Meier") %>%
    select(Estimator, r_0, r_1, rr, RR_lower, RR_upper, rd, RD_lower, RD_upper)
    # mutate(rr = r_1 / r_0,
    #        selnrr = sqrt((1/r_0)^2 * se_0^2 + (1/r_1)^2*se_1^2),
    #        lcl = exp(log(rr) - 1.96*selnrr),
    #        ucl = exp(log(rr) + 1.96*selnrr)) %>% 
    # select(r_0, r_1, rr, lcl, ucl)
  
  return(estimate)
  
}

```

### AJ Estimator

Code for function that generates the AJ estimator object.

```{r}

aj_estimator <- function(dataset, outcome_var, outcome_var_t){
  
  #jitter ties
  #set.seed(1234)
  # Not sure if should set seed for jittering outcomes
  dataset2 <- dataset %>% 
    group_by(get(outcome_var_t)) %>% 
    add_tally() %>% 
    ungroup() %>% 
    mutate(
      # Jitter event times
      jitter = runif(nrow(trial2), min = -.01, max = .01), # N in the samp
      time = ifelse(n>1, get(outcome_var_t) + jitter, get(outcome_var_t))
      )
  
  # Run the AJ model
  aj <- survfit(Surv(time, factor(get(outcome_var))) ~ trt, data = dataset2)
  
}

#test <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

```

Code for function that generates RRs from the AJ estimator.

```{r}
#ASK JESS ABOUT t_val specification

aj_rr <- function(aj_object, t_val){
  
  mod <- summary(aj_object)

  summod <- data.frame(t = mod$time,
                       r = mod$pstate[,2], 
                       se = mod$std.err[,2],
                       # This is not right yet
                       #trt = c(rep(0, mod$strata["trt=0"]), rep(1, mod$strata["trt=1"]))
                       trt = c(rep(0, length(mod[["strata"]][mod[["strata"]] == "trt=0"])), 
                                   rep(1, length(mod[["strata"]][mod[["strata"]] == "trt=1"])))
                       ) %>%
    filter(t < t_val+0.5) %>%  # Deal with the jittering of outcomes
    group_by(trt) %>% 
    summarize(r = last(r),
              se = last(se),
              .groups = 'drop') %>% 
    select(r,se,trt) %>% 
      pivot_wider(names_from = trt,
                  values_from = c(r, se),
                  names_glue = "{.value}_{trt}") %>% 
      rowwise() %>% 
      mutate(rr = r_1/r_0,
           rd = r_1 - r_0,
           # use delta method to get 95% CI
           logRR = log(rr),
           var_1 = se_1^2,
           var_0 = se_0^2,
           var_logRR = (1/r_0)^2 * var_0 + (1/r_1)^2 * var_1,
           logRR_lower = logRR - 1.96*sqrt(var_logRR),
           logRR_upper = logRR + 1.96*sqrt(var_logRR),
           RR_lower = exp(logRR_lower),
           RR_upper = exp(logRR_upper),
           se_rd = sqrt(se_1^2 + se_0^2),
           RD_lower = rd - 1.96*se_rd,
           RD_upper = rd + 1.96*se_rd,
           Estimator = "Aalen-Johansen") %>%
    select(Estimator, r_0, r_1, rr, RR_lower, RR_upper, rd, RD_lower, RD_upper)
  
  return(summod)
    
}

# aj_rr(test)

```

Code for function that plots the AJ object. **This still needs some work to be sure that it's outputting everything correctly.** Currently, the x-axis is wrong. Should be from index date.

```{r}

aj_plot <- function(aj_object, xlab="Gestational weeks post-conception", ylab="Probability in State",
                    outcome = "", competing = ""){
  plot(aj_object, col=c(1,2,1,2), lty=c(2,2,1,1),
       mark.time=FALSE, lwd=2,  #xscale=12,
       xlab=xlab, ylab=ylab)
  legend("topleft", legend = c(paste(competing, ":untreated"), paste(competing, ":treated"),
                               paste(outcome, ":untreated"), paste(outcome, ":treated") 
                               ),
         col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

  # legend(240, .6, c(paste(outcome, ":untreated"), paste(outcome, ":treated"), paste(competing, ":untreated"), paste(competing, ":treated")),
  #        col = c(1, 2, 1, 2), lty = c(1, 1, 2, 2), lwd = 2, bty = 'n')

}

#aj_plot(test, outcome = "Composite", competing = "Delivery")


## VERY HELPFUL: https://cran.r-project.org/web/packages/survival/vignettes/compete.pdf

## Plotting AJ estimators:
#https://search.r-project.org/CRAN/refmans/Epi/html/plotCIF.html

## More AJ R code:
#https://cran.r-project.org/web/packages/AalenJohansen/index.html

```

## Summarize Bias

Code for function that summarizes the bias between KM and AJ estimators.

```{r}

compare <- function(km, aj){
  
  # Calculate the differences
  differences <- tibble(
      # Absolute Risks
      r_0 = km$r_0 - aj$r_0,
      r_1 = km$r_1 - aj$r_1,
      # Risk Ratios
      rr = log(km$rr) - log(aj$rr),
      RR_lower = log(km$RR_lower) - log(aj$RR_lower),
      RR_upper = log(km$RR_upper) - log(aj$RR_upper),
      # Risk differences
      rd = km$rd - aj_rrs_composite$rd,
      RD_lower = km$RD_lower - aj$RD_lower,
      RD_upper = km$RD_upper - aj$RD_upper,
      Estimator = "Differences: KM-AJ"
  )
  
  # Make the kable object
  kable_object <- rbind(km, aj) %>% 
    rbind(differences) %>% 
    ungroup() %>% 
    mutate(r_0 = round(r_0, 2),
           r_1 = round(r_1, 2),
           rr = round(rr, 2),
           RR_lower = round(RR_lower, 2),
           RR_upper = round(RR_upper, 2),
           rd = round(rd, 2),
           RD_lower = round(RD_lower, 2),
           RD_upper = round(RD_upper, 2)) %>% 
    #kbl() %>% 
    kable(col.names = c("Estimator", "Risk Among Untreated", "Risk Among Treated","RR", "RR 95% LCL",
                        "RR 95% UCL", "RD", "RD 95% LCL", "RD 95% UCL"),
          align = "lcccccccc") %>% 
    kable_classic() %>% 
    row_spec(3, bold = T, color = "white", background = "black")
  
  return(kable_object)
  
}


```

## Composite Preeclampsia and Fetal Death

### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome - made earlier
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make composite
    composite = case_when(censor == 1 ~ 0,
                          preg_outcome_final == "fetaldeath" & censor == 0 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 1 ~ 1,
                          preg_outcome_final == "livebirth" & censor == 0 & preeclampsia == 0 ~ 0) #,
    # composite = ifelse(preg_outcome_final == "fetaldeath" | preeclampsia == 1 & censor == 0,
    #                         1,
    #                         0),
    # composite_gw = ifelse(censor == 0,
    #                       preg_outcome_final_gw,
    #                       first_censoring_gw)
    )

## Run the KM analysis only considering the composite outcome.
#wrisk <- km_estimator(trial2, "composite","composite_gw")
wrisk <- km_estimator(trial2, "composite","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for Composite Outcome")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, heigh = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

#############################################################################################
## Calculate the RR at different weeks of follow-up; SE is infinity at 42 weeks of follow-up
## -- ASK JESS about this.
## -- Funky thing: The standard errors get substantially larger as the weeks progress.
## -- Likely because there are fewer and fewer individuals at risk at 42 weeks and imputing
## -- more censoring.
## -- Need to check this intuition with Jess.

km_rr <- km_rr_wCI(wrisk, 41)


```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make a multi-level outcome
    composite_multi = case_when(censor == 1 ~ 0,
                                censor == 0 & preg_outcome_final == "fetaldeath" ~ 1,
                                censor == 0 & preeclampsia == 1 ~ 1,
                                censor == 0 & preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 2
                                ),
    # Make a gestational week variable for the composite_multi variable
    composite_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "composite_multi", "composite_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "composite", competing = "delivery")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate risks without censoring

```{r}

trial3 <- trial %>% 
  mutate(composite = case_when(preg_outcome_final == "fetaldeath" ~ 1,
                               preg_outcome_final == "livebirth" & preeclampsia == 1 ~ 1,
                               preg_outcome_final == "livebirth" & preeclampsia == 0 ~ 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(composite == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## SGA with Fetal Death Only as Competing Event


### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make an indicator for sga, incorporating censoring
    sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0)#,
    # Make a time variable for sga, incorporating censoring
    # sga_t = ifelse(sga == 1 & censor == 0,
    #                preg_outcome_final_gw,
    #                first_censoring_gw)
    )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "sga_km","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for SGA")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

# Undefined at 42 weeks
km_rr <- km_rr_wCI(wrisk, 42)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    sga_multi = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 0), 
    # COULD BE 3 - COME BACK
    # Make a gestational week variable for the sga_multi variable
    sga_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "sga_multi", "sga_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "sga", competing = "no sga")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(sga_out = ifelse(sga == 1, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(sga_out == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## SGA with all Competing Events

### KM Estimator

Output the cumulative incidence curves at the end of follow-up.

```{r}

## Create the composite outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    # censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
    #                 1,
    #                 0),
    # Make an indicator for sga, incorporating censoring
    sga_km = ifelse(sga == 1 & censor == 0,
                    1,
                    0)#,
    # Make a time variable for sga, incorporating censoring
    # sga_t = ifelse(sga == 1 & censor == 0,
    #                preg_outcome_final_gw,
    #                first_censoring_gw)
    )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "sga_km","time")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for SGA")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

# Undefined at 42 weeks
km_rr <- km_rr_wCI(wrisk, 42)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    sga_multi = case_when(censor == 1 ~ 0,
                          censor == 0 & sga == 1 ~ 1,
                          censor == 0 & preg_outcome_final == "fetaldeath" ~ 2,
                          censor == 0 & sga == 0 & preg_outcome_final == "livebirth" ~ 2), 
    # COULD BE 3 - COME BACK
    # Make a gestational week variable for the sga_multi variable
    sga_multi_gw = ifelse(censor == 1,
                                first_censoring_gw,
                                preg_outcome_final_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "sga_multi", "sga_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 42)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "sga", competing = "no sga")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(sga_out = ifelse(sga == 1, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(sga_out == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```

## Preterm live birth

### KM Estimator

Output the cumulative incidence curves at the end of follow-up. Note that we generated gestational weeks from conception, not LMP. Preterm birth is defined as birth <37 weeks of gestation from LMP (2 weeks prior to conception). Thus, we define preterm birth here as birth that occurs at <35 weeks of gestation from conception.

```{r}

## Create the preterm live birth outcome
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & 
                      !is.na(first_censoring_gw),
                    1,
                    0),
    # Make an indicator for sga, incorporating censoring
    preterm_km = ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
                          preg_outcome_final == "livebirth", # Previously 37
                        1,
                        0),
    preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
                       ifelse(preterm_km == 0 & time >= 35, 34,
                              time))
    # preterm_t = ifelse(preterm_km == 1, preg_outcome_final_gw,
    #                ifelse(preg_outcome_final_gw < 35 & preg_outcome_final == "fetaldeath" & 
    #                         censor == 0, preg_outcome_final_gw,
    #                       ifelse(preg_outcome_final_gw >= 35 & censor == 0, 34,
    #                              ifelse(censor == 1 & first_censoring_gw >= 35, 34,
    #                                     ifelse(censor == 1 & first_censoring_gw < 35, 
    #                                            first_censoring_gw, NA)))))

    # preterm_t = case_when(
    #   preterm_km == 1 ~ preg_outcome_final_gw,
    #   preg_outcome_final_gw < 35 & preg_outcome_final == "fetaldeath" & censor == 0 ~ preg_outcome_final_gw,
    #   preg_outcome_final_gw >= 35 & censor == 0 ~ 35,
    #   censor == 1 & first_censoring_gw >= 35 ~ 35,
    #   censor == 1 & first_censoring_gw < 35 ~ first_censoring_gw
    # )

  )


      
  #     ifelse(preg_outcome_final_gw < 35 & censor == 0 & 
  #                        preg_outcome_final == "livebirth", # Previously 37
  #                      preg_outcome_final_gw,
  #                      first_censoring_gw)
  #   ) %>% 
  # # Now, account for the fact that follow-up should stop at 36 weeks.
  # mutate(
  #   # Reset outcome to what it should be at 36 weeks
  #   # This functionally shouldn't matter since preterm can only happen before 36 weeks
  #   preterm_km = ifelse(preterm_t >= 35,
  #                       0,
  #                       preterm_km),
  #   # Reset timing so that it ends at 36 weeks of follow-up
  #   preterm_t = ifelse(preterm_t >= 35,
  #                      35,
  #                      preterm_t)
  # )

## Run the KM analysis only considering the composite outcome.
wrisk <- km_estimator(trial2, "preterm_km","preterm_t")

riskplot <- km_plot(wrisk, "KM Estimator of Risk for Preterm Live Birth")

riskplot

#ggsave(riskplot, file = "q13.png", width = 5, height = 5, units = "cm")

```

Output the risk ratio at the end of follow-up.

```{r}

km_rr <- km_rr_wCI(wrisk, 36)

# km_rr42

```

### AJ Estimator

```{r}

## Make dataset the way that we want it
trial2 <- trial %>% 
  mutate(
    # Make an indicator that the patient was censored before we observed their outcome
    censor = ifelse(first_censoring_gw < preg_outcome_final_gw & !is.na(first_censoring_gw),
                    1,
                    0),
    # Make a multi-level outcome
    preterm_multi = case_when(censor == 1 ~ 0,
                              censor == 0 & preg_outcome_final == "livebirth" & 
                                preg_outcome_final_gw < 35 ~ 1,
                              censor == 0 & preg_outcome_final == "livebirth" &
                                preg_outcome_final_gw >= 35 ~ 2,
                              censor == 0 & preg_outcome_final == "fetaldeath" ~ 2),
    # Don't think we would calculate risks after 36 weeks since no longer at risk. Fetal 
    # -- death is only competing event.
    # However,  do need to leave them in if not limiting all of the data to 36 weeks. 
    # -- Need to clean up these data more.
    # Make a gestational week variable for the sga_multi variable
    preterm_multi_gw = ifelse(censor == 1,
                              first_censoring_gw,
                              preg_outcome_final_gw)
  ) %>% 
  #Now, account for the fact that follow-up should stop at 36 weeks
  mutate(
    # Fix the outcomes to no event (i.e., censor) if the event timing is after 36 weeks
    preterm_multi = ifelse(preterm_multi_gw >= 35,
                           0, # i.e., treat as censoring event.
                          # 2, #i.e., safe from preterm birth, even if ultimatley end in fetal death
                           preterm_multi),
    # Cap the event timing at 36 weeks
    preterm_multi_gw = ifelse(preterm_multi_gw >= 35,
                              36, # i.e., if treat as censoring event
                              #36,
                              preterm_multi_gw)
  )


## Run the competing events analysis

# Make the AJ object

aj <- aj_estimator(trial2, "preterm_multi", "preterm_multi_gw")

# Calculate the RRs

aj_rrs_composite <- aj_rr(aj, 36)

# aj_rrs_composite

# Print the plot

aj_plot(aj_object = aj, outcome = "preterm", competing = "fetal death")

```

### Estimate the bias between KM and RR

```{r}

compare(km_rr, aj_rrs_composite)

```

### Estimate without censoring

### Estimate without censoring

```{r}

trial3 <- trial %>% 
  mutate(preterm = ifelse(preg_outcome_final == "livebirth" & preg_outcome_final_gw < 35, 1, 0))

trial3 %>% 
  group_by(trt) %>% 
  summarize(risk = sum(preterm == 1) / n()) %>% 
  pivot_wider(names_from = trt,
              values_from = c(risk),
              names_glue = "{.value}_{trt}") %>% 
  rowwise() %>% 
  mutate(rr = risk_1 / risk_0,
         rd = risk_1 - risk_0) %>% 
  kable() %>% 
  kable_classic()

```