---
title: "Descriptive Results of Mega Population"
author: "Chase Latour"
date: "1/29/2024"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Call in the libraries that we need

library(kableExtra)
library(tidyverse)

```

# Overview

This HTML file provides the results for our pregnancy simulation focusing on modeling late-pregnancy outcomes as competing events. Pregnancies were only included in the clinical trial if they reached their 4th week of gestation from conception (6 from LMP).

# Set Up

Call in the necessary libraries, files, etc. that we will need.

```{r hide=TRUE}

# Call in the data that we need
s1 <- readRDS("describe trial_scenario1.rds") %>% 
  mutate(Scenario = "Scenario 1")
s2 <- readRDS("describe trial_scenario2.rds") %>% 
  mutate(Scenario = "Scenario 2")
s3 <- readRDS("describe trial_scenario3.rds") %>% 
  mutate(Scenario = "Scenario 3")
s4 <- readRDS("describe trial_scenario4.rds") %>% 
  mutate(Scenario = "Scenario 4")
s5 <- readRDS("describe trial_scenario5.rds") %>% 
  mutate(Scenario = "Scenario 5")
s6 <- readRDS("describe trial_scenario6.rds") %>% 
  mutate(Scenario = "Scenario 6")
s7 <- readRDS("describe trial_scenario7.rds") %>% 
  mutate(Scenario = "Scenario 7")
s8 <- readRDS("describe trial_scenario8.rds") %>% 
  mutate(Scenario = "Scenario 8")
s9 <- readRDS("describe trial_scenario9.rds") %>% 
  mutate(Scenario = "Scenario 9")
s10 <- readRDS("describe trial_scenario10.rds") %>% 
  mutate(Scenario = "Scenario 10")
s11 <- readRDS("describe trial_scenario11.rds") %>% 
  mutate(Scenario = "Scenario 11")
s12 <- readRDS("describe trial_scenario12.rds") %>% 
  mutate(Scenario = "Scenario 12")

# List of data frames
dfs <- mget(paste0("s", 1:12))

# Combine data frames using do.call and rbind
combined_df <- do.call(rbind, dfs)

# Function to replace the prefix for the AJ estimates for the composite outcome
replace_prefix <- function(df, old_prefix, new_prefix) {
  names(df) <- gsub(paste0("^", old_prefix), new_prefix, names(df))
  return(df)
}

combined_df <- replace_prefix(combined_df, "composite_multi_gw", "aj_composite_")

```

# Results

## Clean the Dataset

The code below cleans the dataset for easy review.

```{r}


# Clean the dataset for printing
# - Mutate the variables as desired
# - Pivot the dataset wider

# Calculate the Number of people after censoring with the percentages

censor <- combined_df %>% 
  mutate(n_people_cen = n_people - n_censor) %>% 
  select(Scenario, trt, n_people_cen) %>% 
  group_by(Scenario) %>% 
  summarize(percent_n_cen = round((n_people_cen / sum(n_people_cen))*100, 2),
            n_people_cen = n_people_cen,
            trt = trt,
            .groups = "drop")

wider <- left_join(combined_df, censor, by = c("Scenario","trt")) %>% 
  mutate(n_people = paste0(n_people, " (", percent_n_by_trt, "%)"),
         n_people_cen = paste0(n_people_cen, " (", percent_n_cen, "%)"),
         n_censor = paste0(n_censor, " (", percent_censor, "%)"),
         n_fd_no_censor = paste0(n_fd_no_censor, " (", fd_percent_no_censor, "%)"),
         n_lb_no_censor = paste0(n_lb_no_censor, " (", lb_percent_no_censor, "%)"),
         n_fd_censor = paste0(n_fd_censor, " (", fd_percent_censor, "%)"),
         n_lb_censor = paste0(n_lb_censor, " (", lb_percent_censor, "%)"),
         n_preeclampsia_no_cen = paste0(n_preeclampsia_no_cen, " (", preec_perc_no_cen, "%)"),
         n_sga_no_cen = paste0(n_sga_no_cen, " (", sga_perc_no_cen, "%)"),
         n_preeclampsia_cen = paste0(n_preeclampsia_cen, " (", preec_perc_cen, "%)"),
         n_sga_cen = paste0(n_sga_cen, " (", sga_perc_cen, "%)")
         ) %>% 
  pivot_wider(id_cols = Scenario, names_from = trt, 
                     values_from = c(n_people, n_people_cen, n_censor, n_fd_no_censor, n_lb_no_censor,
                                     n_fd_censor, n_lb_censor, n_preeclampsia_no_cen, 
                                     n_sga_no_cen, n_preeclampsia_cen, n_sga_cen))

```

## Dataset Without Censoring

Below are the descriptive statistics for the cohort at baseline, including those that were censored in analyses.

```{r}

wider %>% 
  select(Scenario, n_people_0, n_people_1, n_censor_0, n_censor_1, n_fd_no_censor_0, n_fd_no_censor_1, 
         n_lb_no_censor_0, n_lb_no_censor_1) %>% 
  knitr::kable(digits = 3, format.args = list(big.mark = ",", scientific = FALSE), col.names = c()) %>% # Trying to add commas to larger numbers
  kable_classic() %>% 
  add_header_above(c("Scenario" = 1, "Untreated" = 1, "Treated" = 1, "Untreated" = 1, "Treated" = 1,
                     "Untreated" = 1, "Treated" = 1, "Untreated" = 1, "Treated" = 1)) %>% 
  add_header_above(c(" " = 1, "Number of Pregnancies" = 2, "Number of Censored Pregnancies" = 2,
                     "Number of Fetal Deaths" = 2, "Number of Live Births" = 2)) 


# From online resource
# add commas to big numbers
# knitr::kable(d, digits = 3, format.args = list(big.mark = ",",
#   scientific = FALSE))

```

# Dataset With Censoring

Below are the descriptive statistics for the cohort at baseline, only considering those that were not censored over follow-up.


```{r}


wider %>% 
  select(Scenario, n_people_cen_0, n_people_cen_1, n_fd_censor_0, n_fd_censor_1, 
         n_lb_censor_0, n_lb_censor_1) %>% 
  knitr::kable(digits = 3, format.args = list(big.mark = ",", scientific = FALSE), col.names = c()) %>% # Trying to add commas to larger numbers
  kable_classic() %>% 
  add_header_above(c("Scenario" = 1, "Untreated" = 1, "Treated" = 1, "Untreated" = 1, "Treated" = 1,
                     "Untreated" = 1, "Treated" = 1)) %>% 
  add_header_above(c(" " = 1, "Number of Pregnancies" = 2, 
                     "Number of Fetal Deaths" = 2, "Number of Live Births" = 2)) 

```

