---
title: "review results_1sim"
author: "Chase Latour"
date: "1/29/2024"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Call in the libraries that we need

library(kableExtra)
library(tidyverse)

```

# Overview

This HTML file provides the results for our pregnancy simulation focusing on modeling late-pregnancy outcomes as competing events. Pregnancies were only included in the clinical trial if they reached their 4th week of gestation from conception (6 from LMP).

# Set Up

Call in the necessary libraries, files, etc. that we will need.

```{r hide=TRUE}

# Call in the data that we need
s1 <- readRDS("analyses_scenario1.rds") %>% 
  mutate(Scenario = "Scenario 1")
s2 <- readRDS("analyses_scenario2.rds") %>% 
  mutate(Scenario = "Scenario 2")
s3 <- readRDS("analyses_scenario3.rds") %>% 
  mutate(Scenario = "Scenario 3")
s4 <- readRDS("analyses_scenario4.rds") %>% 
  mutate(Scenario = "Scenario 4")
s5 <- readRDS("analyses_scenario5.rds") %>% 
  mutate(Scenario = "Scenario 5")
s6 <- readRDS("analyses_scenario6.rds") %>% 
  mutate(Scenario = "Scenario 6")
s7 <- readRDS("analyses_scenario7.rds") %>% 
  mutate(Scenario = "Scenario 7")
s8 <- readRDS("analyses_scenario8.rds") %>% 
  mutate(Scenario = "Scenario 8")
s9 <- readRDS("analyses_scenario9.rds") %>% 
  mutate(Scenario = "Scenario 9")
s10 <- readRDS("analyses_scenario10.rds") %>% 
  mutate(Scenario = "Scenario 10")
s11 <- readRDS("analyses_scenario11.rds") %>% 
  mutate(Scenario = "Scenario 11")
s12 <- readRDS("analyses_scenario12.rds") %>% 
  mutate(Scenario = "Scenario 12")

# List of data frames
dfs <- mget(paste0("s", 1:12))

# Combine data frames using do.call and rbind
combined_df <- do.call(rbind, dfs)

# Function to replace the prefix for the AJ estimates for the composite outcome
replace_prefix <- function(df, old_prefix, new_prefix) {
  names(df) <- gsub(paste0("^", old_prefix), new_prefix, names(df))
  return(df)
}

combined_df <- replace_prefix(combined_df, "composite_multi_gw", "aj_composite_")

```

# Results

## Functions

We will use these functions to print out the results below.

```{r}


# Clean up the dataset according to outcome

transform <- function(dset, var_name) {

  # Transformations
  dset2 <- dset %>%
    mutate(
      # Results from KM Estimator
      km_r0 = round(!!sym(paste0("km_", var_name, "_ r_0")), 2),
      km_r1 = round(!!sym(paste0("km_", var_name, "_ r_1")), 2),
      km_rd = round(!!sym(paste0("km_", var_name, "_ rd")) * 100, 2),
      km_rd_ci = paste0("(", round(!!sym(paste0("km_", var_name, "_ rd_lcl"))*100,2), ", ",
                        round(!!sym(paste0("km_", var_name, "_ rd_ucl"))*100,2), ")"),
      
      # Results from AJ Estimator
      aj_r0 = round(!!sym(paste0("aj_", var_name, "_ r_0")), 2),
      aj_r1 = round(!!sym(paste0("aj_", var_name, "_ r_1")), 2),
      aj_rd = round(!!sym(paste0("aj_", var_name, "_ rd")) * 100, 2),
      aj_rd_ci = paste0("(", round(!!sym(paste0("aj_", var_name, "_ rd_lcl"))*100,2), ", ",
                        round(!!sym(paste0("aj_", var_name, "_ rd_ucl"))*100, 2), ")"),
      
      # Results without censoring
      no_censor_r0 = round(!!sym(paste0("no_censor_", var_name, "_ r_0")), 2),
      no_censor_r1 = round(!!sym(paste0("no_censor_", var_name, "_ r_1")), 2),
      no_censor_rd = round(!!sym(paste0("no_censor_", var_name, "_ rd")) * 100, 2),
      no_censor_rd_ci = paste0("(", 
                               round(!!sym(paste0("no_censor_", var_name, "_ rd_lcl"))*100, 2),
                               ", ",
                               round(!!sym(paste0("no_censor_", var_name, "_ rd_lcl"))*100, 2),
                               ")")#,
      
      
      
    ) %>% 
    rowwise() %>% 
    mutate(
      # Truth
      truth_r0 = ifelse(var_name == "composite",
                        round(!!sym(paste0("potential_ risk_", var_name, "_untrt")), 2),
                        round(`potential_ risk_sga_untrt`, 2)),
      truth_r1 = ifelse(var_name == "composite",
                        round(!!sym(paste0("potential_ risk_", var_name, "_trt")), 2),
                        round(`potential_ risk_sga_trt`, 2)),
      truth_rd = ifelse(var_name == "composite",
                        round(!!sym(paste0("potential_ rd_", var_name)) * 100, 2),
                        round(`potential_ rd_sga`, 2))
    ) %>% 
    ungroup() %>% 
    select(c(Scenario, km_r0, km_r1, km_rd, km_rd_ci,
             aj_r0, aj_r1, aj_rd, aj_rd_ci,
             no_censor_r0, no_censor_r1, no_censor_rd, no_censor_rd_ci,
             truth_r0, truth_r1, truth_rd)) %>%
    kable(col.names = c("Scenario", "Risk|Untrt", "Risk|Trt", "RD", "RD 95% CI",
                        "Risk|Untrt", "Risk|Trt", "RD", "RD 95% CI",
                        "Risk|Untrt", "Risk|Trt", "RD", "RD 95% CI",
                        "Risk|Untrt", "Risk|Trt", "RD")) %>%
    kable_classic() %>%
    add_header_above(c(" " = 1, "Kaplan-Meier" = 4, "Aalen-Johansen" = 4, "No Censoring" = 4,
                       "Truth" = 3))
  
  return(dset2)
}

```

## Composite Outcome Results

Now, we will print out those results for the composite outcome


```{r}

## Print out the kable 

transform(combined_df, "composite")
  

```


## SGA: All Competing Events

```{r}

# Print out the kable

transform(combined_df, "sga_all")


```


## SGA: Fetal Death Only as Competing Event


```{r}

transform(combined_df, "sga_fd")

```

