---
title: "review results_1sim"
author: "Chase Latour"
date: "1/29/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This HTML file provides the results for our pregnancy simulation focusing on modeling late-pregnancy outcomes as competing events. The original cohort for each of these was 1,000,000 conceptions. Pregnancies were only included in the clinical trial if they reached their 4th week of gestation from conception (6 from LMP).

# Set Up

Call in the necessary libraries, files, etc. that we will need.

```{r}

# Call in the libraries that we need

library(kableExtra)
library(tidyverse)

# Call in the data that we need
s1 <- readRDS("analyses_scenario1.rds") %>% 
  mutate(Scenario = "Scenario 1")
s2 <- readRDS("analyses_scenario2.rds") %>% 
  mutate(Scenario = "Scenario 2")
s3 <- readRDS("analyses_scenario3.rds") %>% 
  mutate(Scenario = "Scenario 3")
s4 <- readRDS("analyses_scenario4.rds") %>% 
  mutate(Scenario = "Scenario 4")
s5 <- readRDS("analyses_scenario5.rds") %>% 
  mutate(Scenario = "Scenario 5")
s6 <- readRDS("analyses_scenario6.rds") %>% 
  mutate(Scenario = "Scenario 6")
s7 <- readRDS("analyses_scenario7.rds") %>% 
  mutate(Scenario = "Scenario 7")
s8 <- readRDS("analyses_scenario8.rds") %>% 
  mutate(Scenario = "Scenario 8")
s9 <- readRDS("analyses_scenario9.rds") %>% 
  mutate(Scenario = "Scenario 9")
s10 <- readRDS("analyses_scenario10.rds") %>% 
  mutate(Scenario = "Scenario 10")
### These aren't correct right now - read in the wrong data when ran
s11 <- readRDS("analyses_scenario11.rds") %>% 
  mutate(Scenario = "Scenario 11")
s12 <- readRDS("analyses_scenario12.rds") %>% 
  mutate(Scenario = "Scenario 12")

# List of data frames
dfs <- mget(paste0("s", 1:12))

# Combine data frames using do.call and rbind
combined_df <- do.call(rbind, dfs)

```

# Results

## Functions

We will use these functions to print out the results below.

```{r}

# Clean up the dataset according to outcome

transform <- function(dset, var_name) {

  # Transformations
  dset2 <- dset %>%
    mutate(
      km_r0 = round(!!sym(paste0("km_", var_name, "_ r_0")), 2),
      km_r1 = round(!!sym(paste0("km_", var_name, "_ r_1")), 2),
      km_rd = round(!!sym(paste0("km_", var_name, "_ rd")) * 100, 2),
      km_rd_ci = paste0("(", round(!!sym(paste0("km_", var_name, "_ rd_lcl"))*100,2), ", ",
                        round(!!sym(paste0("km_", var_name, "_ rd_ucl"))*100,2), ")"),
      aj_r0 = round(!!sym(paste0("aj_", var_name, "_ r_0")), 2),
      aj_r1 = round(!!sym(paste0("aj_", var_name, "_ r_1")), 2),
      aj_rd = round(!!sym(paste0("aj_", var_name, "_ rd")) * 100, 2),
      aj_rd_ci = paste0("(", round(!!sym(paste0("aj_", var_name, "_ rd_lcl"))*100,2), ", ",
                        round(!!sym(paste0("aj_", var_name, "_ rd_ucl"))*100, 2), ")"),
      truth_r0 = round(!!sym(paste0("no_censor_", var_name, "_ r_0")), 2),
      truth_r1 = round(!!sym(paste0("no_censor_", var_name, "_ r_1")), 2),
      truth_rd = round(!!sym(paste0("no_censor_", var_name, "_ rd")) * 100, 2)
    ) %>% 
    select(c(Scenario, km_r0, km_r1, km_rd, km_rd_ci, aj_r0, aj_r1, aj_rd, aj_rd_ci, truth_r0, truth_r1, truth_rd)) %>% 
    kable(col.names = c("Scenario", "Risk|Untrt", "Risk|Trt", "RD", "RD 95% CI",
                        "Risk|Untrt", "Risk|Trt", "RD", "RD 95% CI",
                        "Risk|Untrt", "Risk|Trt", "RD")) %>% 
    kable_classic() %>% 
    add_header_above(c(" " = 1, "Kaplan-Meier" = 4, "Aalen-Johansen" = 4, "Truth" = 3))
  
  return(dset2)
}

```

## Composite Outcome Results

Now, we will print out those results for the composite outcome



```{r}

## Print out the kable 

transform(combined_df, "composite")

```


## SGA: All Competing Events

```{r}

# Print out the kable

transform(combined_df, "sga_all")


```


## SGA: Fetal Death Only as Competing Event


```{r}

transform(combined_df, "sga_fd")

```

